Improving TensorFlow's Memory Swapping by Prioritizing Earlier Tensors

### 2019-08-18 paper.tex
Move IEEEtran.cls up
Merge abstract
Revise experiment analysis

\documentclass[conference]{classes/IEEEtran}

@@

\begin{abstract}
    The ever-increasing sizes of deep learning models and datasets used increase the need for memory, as insufficient memory may abort a training process. Adding to this problem, deep learning trainings tend to use GPUs over CPUs for better training speed, where in general a GPU has significantly less memory than its CPU counterpart.

    One solution is \textit{memory swapping}, a way to free some memory by temporarily moving data to another memory pool, in this case moving data from a GPU memory to a CPU memory. However, since moving data takes time---the larger the data the more the time---performing memory swapping in a training can significantly increase the training duration. Therefore, an ideal memory swapping has to be selective on how much and which data to swap.

    TensorFlow, a machine learning framework, features memory swapping that, based on our analysis over its implementation, can be improved to reduce the increase on training duration it causes. The improvement is done is by prioritizing earlier tensors (\textit{tensor} is the basic data unit in TensorFlow) which, because of a certain backpropagation property, allows the asynchronicity of program execution to increase, ultimately reducing training duration.

    Based on our experiments on Char-RNN models with various hyperparameters and datasets (of size 285 KB and 4.4 MB), the improvement reduces up to around 3\% of training duration on certain cases.
\end{abstract}

@@

The experiment cases can be divided into two parts, the ones using the small dataset and the ones using the large dataset. The small dataset is a text containing 291,268 characters (285 KB) while the large dataset is a text containing 4,573,338 characters (4.4 MB). We deem the contents of the text unimportant for the paper's context, but between both datasets, the contents are different.

@@

The experiment results are shown in \autoref{tbl:Overview}, where each case is respective to the one having the same ID in \autoref{tbl:Cases}.

@@

In all cases, means of both versions are statistically different according to t-test, except for cases A1, A2, A3, and B1. The similarity of those four cases is that they use low timestep. The other two cases with low timestep (B2 and B3) also yield low improvements. Meanwhile, all cases with medium and high timesteps yield significant improvements. These reasons lead us to conclude that the timestep need to exceed a certain threshold to yield significant improvements.

>>>>

\documentclass[conference]{IEEEtran}

@@

\begin{abstract}
    The ever-increasing sizes of deep learning models and datasets used increase the need for memory, as insufficient memory may abort a training process. Adding to this problem, deep learning trainings tend to use GPUs over CPUs for better training speed, where in general a GPU has significantly less memory than its CPU counterpart.
    One solution is \textit{memory swapping}, a way to free some memory by temporarily moving data to another memory pool, in this case moving data from a GPU memory to a CPU memory. However, since moving data takes time---the larger the data the more the time---performing memory swapping in a training can significantly increase the training duration. Therefore, an ideal memory swapping has to be selective on how much and which data to swap.
    TensorFlow, a machine learning framework, features memory swapping that, based on our analysis over its implementation, can be improved to reduce the increase on training duration it causes. The improvement is done is by prioritizing earlier tensors (\textit{tensor} is the basic data unit in TensorFlow) which, because of a certain backpropagation property, allows the asynchronicity of program execution to increase, ultimately reducing training duration.
    Based on our experiments on Char-RNN models with various hyperparameters and datasets (of size 285 KB and 4.4 MB), the improvement reduces up to around 3\% of training duration on certain cases.
\end{abstract}

@@

The experiment cases can be divided into two parts, the ones using the small dataset and the ones using the large dataset. The small dataset is a text containing 291,268 characters (285 KB) while the large dataset is a text containing 4,573,338 characters (4.4 MB). We deem the contents of the text unimportant for the paper's purpose, but between both datasets, the contents are different.

@@

\autoref{tbl:Overview} shows the means of execution times of all cases, with IDs respective to the ones in \autoref{tbl:Cases}.

@@

In all cases, means of both versions are statistically different according to t-test, except for cases A1, A2, A3, and B1. The similarity of those four cases is that they use a low timestep. The other two cases with low timestep (B2 and B3) also yield low improvements. Meanwhile, all cases with medium and high timesteps yield significant improvements. These reasons lead us to conclude that the timestep need to exceed a certain threshold to yield significant improvements. To make it clearer, even though case A1 yields a negative result, means of both versions in the case are not statistically different, and therefore not a concern to us.
